---
project:
  type: website

website:  
  title: "Occupación tigrinus"
  subtitle: "co-ocurrencia tigrinus y perros"
  author: 
  - name: Alejandra Bonilla
    orcid: https://orcid.org/0000-0003-0803-5884
  - name: Diego J. Lizcano
    orcid: https://orcid.org/0000-0002-9648-0576
toc: true
license: CC BY-SA
page-navigation: true
navbar:
    right:
      - icon: github
        href: https://github.com/dlizcano/tigrinus_occu
        aria-label: GitHub

execute: 
  echo: true
  eval: false
  warning: false
  error: false
  output: asis

citation: true
google-scholar: true
bibliography: grateful-refs.bib
bibliographystyle: https://www.zotero.org/styles/tapir-conservation
  
format: 
  html: 
    code-fold: true
    code-summary: "codigo R"
    code-block-bg: true
    highlight-style: breezedark
    code-block-border-left: "#31BAE9"
    fig-width: 8
    fig-height: 6
---

Multispecies occupancy model from Rota et al. (2016), for two (or more) interacting species. The model generalizes the standard single-species occupancy model from MacKenzie et al. (2002), as a multispecies occupancy model designed for analyzing presence/absence data of two or more interacting species. It builds upon the standard single-species occupancy model by incorporating the potential for interactions between species. The model estimates occupancy probabilities for each species (Marginal occupancy), as well as the strength and direction of interactions between them (contitional occupancy), and can also incorporate the effects of environmental covariates.

On the interpretation of interaction term: https://groups.google.com/g/unmarked/c/W3xceS-hSNA/m/8OvCD93BAQAJ

If the estimate of species1:species2 intercept is positive, then species 1 has greater occupancy probability at sites occupied by species 2. The reverse is also true: species 2 has greater occupancy at sites occupied by species 1. It might be more helpful to think about the value of the parameter it this way: are species 1 and 2 likely to occupy the same sites (species1:species2 positive), different sites (species1:species2 negative), or be independent of each other (species1:species2 = 0)?

We used *Leopardus tigrinus* and domestic dog data from 8 regions across the Colombian Andes, sampled with arrays of 20 to 30 camera traps, to exploring evidence for interactions with domestic dog while accounting for the effects of environmental variables.

# cargar paquetes

```{r}

library(knitr)
library(mapview) # mapas facil
library(readxl) #leer datos
library(xlsx) # write dato
library(readr) # lee datos
library(sf) # vector map
library(geodata) # replace getData de raster para Terra
library(raster) # mapas raster
library(spatstat) # interpola mapa
# library(maptools) # to coerce to ppp. note that 'maptools' will be retired by the end of 2023
# library(rgdal) # rgdal will be retired during 2023 #some tricks to change projection
# library(stars)
# library(unmarked) # occu models
library(DT) # html table
library(camtrapR) # camera trap data creation
library(terra) # new raster
library(elevatr) # get elevation
library(DT) # tables
library (tmap) # nice maps
library(rnaturalearth) # colombia map

library(grateful) # citation packages
library(tidyverse) # maneja datos

# source("C:/CodigoR/tigrinus2/R/organizadato.R") # old version
source("C:/CodigoR/WCS-CameraTrap/R/organiza_datos_v3.R") # new version

```

## Cargar datos

### Ucumari, Pitalito, La cocha1, La Fe, Rio Grande, La cocha2, Ituango, Chingaza y Saldaña.

```{r}


Full_data_ucu <- read_excel("D:/BoxFiles/Box Sync/CodigoR/tigrinus/data/Full_data_Ucumari_Huila_Cocha1_Cocha2.xlsx", 
    sheet = "ucumari", col_types = c("numeric", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "numeric", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "numeric", 
        "numeric", "numeric", "numeric", 
        "text"))

Full_data_pitalito <- read_csv("D:/BoxFiles/Box Sync/CodigoR/tigrinus/data/huila_merged.csv", 
    col_types = cols(`Date_Time Captured` = col_character(), 
        camera_trap_start_date = col_character(), 
        camera_trap_end_date = col_character()))


Full_data_pitalito <- read_csv("D:/BoxFiles/Box Sync/CodigoR/tigrinus/data/huila_merged.csv", 
    col_types = cols(`Date_Time Captured` = col_character(), 
        camera_trap_start_date = col_character(), 
        camera_trap_end_date = col_character()))

Full_data_cocha1  <- read_csv("D:/BoxFiles/Box Sync/CodigoR/tigrinus/data/Cocha1_merged.csv", 
    col_types = cols(camera_trap_start_date = col_character(), 
        camera_trap_end_date = col_character()))



Full_data_cocha2  <- read_csv("D:/BoxFiles/Box Sync/CodigoR/tigrinus/data/Cocha_2.csv", 
    col_types = cols(camera_trap_start_date = col_character(), 
                     "Photo time" = col_character(),
                     "Photo Date" = col_character(),
        camera_trap_end_date = col_character())) 

Full_data_cocha2$camera_trap <- Full_data_cocha2$`Camera Trap Name`


lafe_data <- read_excel("C:/CodigoR/tigrinus2/data/cuencaverde.xlsx",   sheet = "LaFe_2021")

lafe_data$year <- year(lafe_data$Photo_Date)

# filter by 2021
# lafe_data <- lafe_data_raw %>% filter(year == "2021")

riogrande_data <-  read_excel("C:/CodigoR/tigrinus2/data/cuencaverde.xlsx", 
    sheet = "Riogrande_completo") |> mutate(yr=year(Photo_Date)) |> filter(yr==2021) 

ituango_data <-  read_excel("D:/BoxFiles/Box Sync/CodigoR/tigrinus3/Ituango/Base de  Datos Final.xlsx", 
    sheet = "Datos_camaras12-07-2015") |> mutate(yr=Year) |> filter(yr==2015) |> filter (MES>=3 & MES<=6)


# Dista Casa Ituango en ituango_cov_occ
ituango_dis_casa <- read_csv("C:/CodigoR/tigrinus2/data/ituango_cov_occ.csv")




chingaza <- read_excel("D:/BoxFiles/Box Sync/CodigoR/tigrinus3/Chingaza/tigrinus_Perros_PNNChingaza.xlsx") |>
  mutate(scientificName = paste(Genus, Species, sep= " "))  
chingaza$year <-  lubridate::year(chingaza$Photo_date)
chingaza_data <- chingaza |> filter(chingaza$Photo_date <= "2020-01-31")

#### usar Det_Hist_diario_cf_Ch_1year_.csv Det_Hist_diario_Lt_Ch_1year_.csv historias deteccion de  ituango



perro_Saldania <- read_csv("C:/CodigoR/tigrinus2/data/perro_Saldania.csv")
tigrinus_Saldania <- read_csv("C:/CodigoR/tigrinus2/data/tigrinus_Saldania.csv")



```

#### Get elevation and distance covariates

## falta: to include forest (GFW) cover y huella humana

```{r}




#casas <- read_csv("C:/CodigoR/tigrinus2/data/casas.csv")
# casas_sf <- st_as_sf(casas, coords = c("lon", "lat"), crs = "EPSG:4326")
casas_sf <- st_read("C:/CodigoR/tigrinus2/data/casas.shp")



############# start spatial part
#### make sf object
ucumari <- Full_data_ucu |> 
  select("camera_trap",
         "Latitude",
         "Longitude",
         "camera_trap") |>
  dplyr::distinct( ) |> 
  mutate(region="Ucumari RP")
  

pitalito <- Full_data_pitalito |> 
  select("camera_trap",
         "Latitude",
         "Longitude",
         "camera_trap") |>
  dplyr::distinct( ) |> 
  mutate(region="Pitalito")

cocha1 <- Full_data_cocha1 |> 
  select("camera_trap",
         "Latitude",
         "Longitude",
         "camera_trap") |>
  dplyr::distinct( ) |> 
  mutate(region="Cocha1")

cocha2 <- Full_data_cocha2 |> 
  select("camera_trap",
         "Latitude",
         "Longitude",
         "camera_trap") |>
  dplyr::distinct( ) |>  
  mutate(region="Cocha2")

lafe <- lafe_data  |> 
  select("camera_trap",
         "Latitude",
         "Longitude",
         "camera_trap") |>
  dplyr::distinct( ) |> 
  mutate(region="La Fe")

riogrande <- riogrande_data  |> 
  select("camera_trap",
         "Latitude",
         "Longitude",
         "camera_trap") |>
  dplyr::distinct( ) |> 
  mutate(region="Rio Grande")

ituango <- ituango_data  |> 
  select("camera_trap",
         "Latitude",
         "Longitude",
         "camera_trap") |>
  dplyr::distinct( ) |>  
  mutate(region="Ituango")




chingaza <- chingaza_data  |> 
  select("camera_trap",
         "Latitude",
         "Longitude",
         "camera_trap") |>
  dplyr::distinct( ) |> 
  mutate(region="Chingaza NP")



saldania <- tigrinus_Saldania  |> mutate(camera_trap=deployment_id_cam) |> 
  mutate(Longitude=longitude) |> 
  mutate(Latitude=latitude) |> 
  select("camera_trap",
         "Latitude",
         "Longitude",
         "camera_trap") |>
  dplyr::distinct( ) |> 
  mutate(region="Saldaña")


# join
puntos <- rbind(ucumari,
                pitalito, 
                cocha1, 
                lafe,
                riogrande,
                ituango,
                cocha2,
                chingaza,
                saldania)





################
# plot map
################
# 
# pal = mapviewPalette("mapviewTopoColors") #color palete
# 
# mapview(elev_ucu_ras) + 
#   mapview(casa_dist_rast, col.regions = pal(100), at = seq(600, 240000, 100), legend = TRUE) +
#   mapview(puntos_sf["camera_trap"]) 

############### end spatial part



```

#### crerar historias detección

```{r}

# make species names
Full_data_ucu$binomial <- str_c (Full_data_ucu$Genus, "_", Full_data_ucu$Species)

Full_data_pitalito$binomial <- Full_data_pitalito$`Genus Species`


# #funcion para crear todas las tablas de datos
# all_data_ucu <-  f.matrix.creator2 (Full_data_ucu)
# 
# # names(all_data) # ver lass especies y en que lista esta cada una
# # kable(names(all_data)) # html table
# # Tigrinus es lista 8
# 
# datatable(
# data = as.data.frame(names(all_data_ucu)),
# caption = "Especies Ucumari",
# filter = "top"
# )

# fix date format
# Ucumari

Full_data_ucu$start_date <- as.Date(Full_data_ucu$"camera_trap_start_date", "%Y-%m-%d")
Full_data_ucu$end_date <- as.Date(Full_data_ucu$"camera_trap_end_date", "%Y-%m-%d")
Full_data_ucu$eventDate <- as.Date(Full_data_ucu$"Photo_Date", "%Y-%m-%d")
Full_data_ucu$eventDateTime <- ymd_hms(paste(Full_data_ucu$"Photo_Date", Full_data_ucu$"Photo time", sep=" "))
# rename camera id
Full_data_ucu$camid <- Full_data_ucu$`camera_trap`

# Pitalito
Full_data_pitalito$start_date <- as.Date(Full_data_pitalito$"camera_trap_start_date", "%Y-%m-%d")
Full_data_pitalito$end_date <- as.Date(Full_data_pitalito$"camera_trap_end_date", "%Y-%m-%d")
# Full_data_pitalito$eventDate <- as.Date(Full_data_pitalito$, "%Y-%m-%d")
Full_data_pitalito$eventDateTime <- ymd_hms(Full_data_pitalito$`Date_Time Captured`)
# rename camera id
Full_data_pitalito$camid <- Full_data_pitalito$camera_trap

# La Fe

lafe_data$start_date <- as.Date(lafe_data$"camera_trap_start_date", "%Y-%m-%d")
lafe_data$end_date <- as.Date(lafe_data$"camera_trap_end_date", "%Y-%m-%d")
lafe_data$eventDate <- as.Date(lafe_data$Photo_Date, "%Y-%m-%d")

lafe_data$eventDateTime <- ymd_hms(paste(lafe_data$"Photo_Date", lafe_data$"Photo time", sep=" "))# rename camera id
lafe_data$camid <- lafe_data$camera_trap



# Rio Grande

riogrande_data$start_date <- as.Date(riogrande_data$"camera_trap_start_date", "%Y-%m-%d")
riogrande_data$end_date <- as.Date(riogrande_data$"camera_trap_end_date", "%Y-%m-%d")
riogrande_data$eventDate <- as.Date(riogrande_data$Photo_Date, "%Y-%m-%d")


riogrande_data$eventDateTime <- lubridate::ymd_hm(paste(
  as.Date(riogrande_data$"Photo_Date", "%Y-%m-%d"), 
 riogrande_data$"Photo time", 
  sep=" "),
  tz = "America/Bogota")

# rename camera id
riogrande_data$camid <- riogrande_data$camera_trap


# Ituango add start end
ituango_data$start_date <- as.Date("2015-03-01", "%Y-%m-%d")
ituango_data$end_date <- as.Date("2015-06-30", "%Y-%m-%d")
ituango_data$eventDate <- as.Date(
                        paste(ituango_data$Year, 
                              ituango_data$MES,
                              ituango_data$DIA,
                              sep="-"
                              ),"%Y-%m-%d")


ituango_data$eventDateTime <- lubridate::ymd_hm(paste(
  as.character(ituango_data$eventDate), 
 ituango_data$"HORA", 
  sep=" "),
  tz = "America/Bogota")


# rename camera id
ituango_data$camid <- ituango_data$camera_trap





# Chingaza 
chingaza_data$start_date <- as.Date("2019-09-26", "%Y-%m-%d")# as.Date(chingaza_data$"Camera Start Date", "%Y-%m-%d")
chingaza_data$eventDateTime <- lubridate::ymd_hms(
   paste(
     as.character(chingaza_data$Photo_date), 
     chingaza_data$Photo_time, 
     sep=" ")
   #tz = "America/Bogota",
   )

chingaza_data$end_date <- NA
# loop to fix dates
for (i in 1:length(chingaza_data$"Camera End Date")) {
    if (chingaza_data$"Camera End Date"[i]<="2020-01-31") {
    chingaza_data$end_date[i] = as.Date(chingaza_data$"Camera End Date", "%Y-%m-%d")[i]
    } else {
    chingaza_data$end_date[i] = as.Date("2020-02-01", "%Y-%m-%d")
    }
  # print(chingaza_data$end_date[i])
}
chingaza_data$end_date <- as.Date(chingaza_data$end_date)
#as.Date("2020-09-01", "%Y-%m-%d")
# rename camera id
chingaza_data$camid <- chingaza_data$camera_trap









# filter 2021 and make uniques
ucu_CToperation  <- Full_data_ucu |> dplyr::group_by(camid) |> #(array_locID) |> 
                           mutate(minStart=start_date, maxEnd=end_date) |> distinct(Longitude, Latitude, minStart, maxEnd) |> dplyr::ungroup()
# remove one duplicated
# View(CToperation)
# CToperation <- CToperation[-15,]


pitalito_CToperation  <- Full_data_pitalito |> dplyr::group_by(camid) |> #(array_locID) |> 
                           mutate(minStart=start_date, maxEnd=end_date) |> distinct(Longitude, Latitude, minStart, maxEnd) |> dplyr::ungroup()



lafe_CToperation  <- lafe_data |> dplyr::group_by(camid) |> #(array_locID) |> 
                           mutate(minStart=start_date, maxEnd=end_date) |> distinct(Longitude, Latitude, minStart, maxEnd) |> dplyr::ungroup()


### selecting 2021
riogrande_CToperation  <- riogrande_data |> 
    dplyr::group_by(camid) |> #(array_locID) |> 
                           mutate(minStart=start_date, maxEnd=end_date) |> distinct(Longitude, Latitude, minStart, maxEnd) |> dplyr::ungroup() 


ituango_CToperation  <- ituango_data |> 
    dplyr::group_by(camid) |> #(array_locID) |> 
                           mutate(minStart=start_date, maxEnd=end_date) |> distinct(Longitude, Latitude, minStart, maxEnd) |> dplyr::ungroup() 

chingaza_CToperation  <- chingaza_data |> 
    dplyr::group_by(camid) |> #(array_locID) |> 
                           mutate(minStart=start_date, maxEnd=end_date) |> distinct(Longitude, Latitude, minStart, maxEnd) |> dplyr::ungroup() 



# remove dos problematic in lafe
# lafe_CToperation <- lafe_CToperation[-c(28,29),]

# Generamos la matríz de operación de las cámaras

ucu_camop <- cameraOperation(CTtable= ucu_CToperation, # Tabla de operación
                         stationCol= "camid", # Columna que define la estación
                         setupCol= "minStart", #Columna fecha de colocación
                         retrievalCol= "maxEnd", #Columna fecha de retiro
                         #hasProblems= T, # Hubo fallos de cámaras
                         dateFormat= "%Y-%m-%d")#, #, # Formato de las fechas
                         #cameraCol="Camera_Id")
                         # sessionCol= "Year")


pitalito_camop <- cameraOperation(CTtable= pitalito_CToperation, # Tabla de operación
                         stationCol= "camid", # Columna que define la estación
                         setupCol= "minStart", #Columna fecha de colocación
                         retrievalCol= "maxEnd", #Columna fecha de retiro
                         #hasProblems= T, # Hubo fallos de cámaras
                         dateFormat= "%Y-%m-%d")#, #, # Formato de las fechas
                         #cameraCol="Camera_Id")
                         # sessionCol= "Year")


lafe_camop <- cameraOperation(CTtable= lafe_CToperation, # Tabla de operación
                         stationCol= "camid", # Columna que define la estación
                         setupCol= "minStart", #Columna fecha de colocación
                         retrievalCol= "maxEnd", #Columna fecha de retiro
                         #hasProblems= T, # Hubo fallos de cámaras
                         dateFormat= "%Y-%m-%d")#, #, # Formato de las fechas
                         #cameraCol="Camera_Id")
                         # sessionCol= "Year")





riogrande_camop <- cameraOperation(CTtable= riogrande_CToperation, # Tabla de operación
                         stationCol= "camid", # Columna que define la estación
                         setupCol= "minStart", #Columna fecha de colocación
                         retrievalCol= "maxEnd", #Columna fecha de retiro
                         #hasProblems= T, # Hubo fallos de cámaras
                         dateFormat= "%Y-%m-%d")#, #, # Formato de las fechas
                         #cameraCol="Camera_Id")
                         # sessionCol= "Year")


ituango_camop <- cameraOperation(CTtable= ituango_CToperation, # Tabla de operación
                         stationCol= "camid", # Columna que define la estación
                         setupCol= "minStart", #Columna fecha de colocación
                         retrievalCol= "maxEnd", #Columna fecha de retiro
                         #hasProblems= T, # Hubo fallos de cámaras
                         dateFormat= "%Y-%m-%d")#, #, # Formato de las fechas
                         #cameraCol="Camera_Id")
                         # sessionCol= "Year")


chingaza_camop <- cameraOperation(CTtable= chingaza_CToperation, # Tabla de operación
                         stationCol= "camid", # Columna que define la estación
                         setupCol= "minStart", #Columna fecha de colocación
                         retrievalCol= "maxEnd", #Columna fecha de retiro
                         #hasProblems= T, # Hubo fallos de cámaras
                         dateFormat= "%Y-%m-%d")#, #, # Formato de las fechas
                         #cameraCol="Camera_Id")
                         # sessionCol= "Year")


# Generar las historias de detección ---------------------------------------
## remove problem species

Full_data_ucu$scientificName <- paste(Full_data_ucu$Genus, 
                                      Full_data_ucu$Species, 
                                      sep=" ")

#### remove setups
ucu_ind <- which(Full_data_ucu$scientificName=="NA NA")
Full_data_ucu <- Full_data_ucu[-ucu_ind,]

# ind <- which(Ecu_full$scientificName=="Set up")
# Ecu_full <- Ecu_full[-ind,]
# 
# ind <- which(Ecu_full$scientificName=="Blank")
# Ecu_full <- Ecu_full[-ind,]
# 
# ind <- which(Ecu_full$scientificName=="Unidentifiable")
# Ecu_full <- Ecu_full[-ind,]

Full_data_pitalito$scientificName <- Full_data_pitalito$`Genus Species`

#### remove setups and NAs
pitalito_ind <- which(is.na(Full_data_pitalito$scientificName))
Full_data_pitalito <- Full_data_pitalito[-pitalito_ind,]

# fix lafe
lafe_data$scientificName <- lafe_data$binomial
# fix riogrande
riogrande_data$scientificName <- riogrande_data$binomial

#### remove setups and NAs
riogrande_data_ind <- which(riogrande_data$scientificName=="NA_NA")
riogrande_data <- riogrande_data[-riogrande_data_ind,]

#### remove setups and NAs
ituango_data_ind <- which(ituango_data$scientificName=="NA")
ituango_data <- ituango_data[-ituango_data_ind,]



############### Ucu

ucu_DetHist_list <- lapply(unique(Full_data_ucu$scientificName), FUN = function(x) {
  detectionHistory(
    recordTable         = Full_data_ucu, # tabla de registros
    camOp                = ucu_camop, # Matriz de operación de cámaras
    stationCol           = "camid",
    speciesCol           = "scientificName",
    recordDateTimeCol    = "eventDateTime",
    recordDateTimeFormat  = "%Y-%m-%d %H:%M:%S",
    species              = x,     # la función reemplaza x por cada una de las especies
    occasionLength       = 8, # Colapso de las historias a 10 ías
    day1                 = "survey", # "survey" a specific date, "station", #inicie en la fecha de cada survey
    datesAsOccasionNames = TRUE,
    includeEffort        = TRUE,
    scaleEffort          = FALSE,
    #unmarkedMultFrameInput=TRUE
    timeZone             = "America/Bogota" 
    )
  }
)

# names
names(ucu_DetHist_list) <- unique(Full_data_ucu$scientificName)
# Finalmente creamos una lista nueva donde estén solo las historias de detección
ucumari_ylist <- lapply(ucu_DetHist_list, FUN = function(x) x$detection_history)
# otra lista con effort scaled
ucumari_efort <- lapply(ucu_DetHist_list, FUN = function(x) x$effort) 


my_path <- "C:/CodigoR/tigrinus2/data/ToAleja/"

write.xlsx2(ucumari_ylist[[7]], paste0(my_path, "Ucumari.xlsx"), row.names = FALSE, sheetName = "tigrinus")                                      # Write xlsx file
write.xlsx2(ucumari_ylist[[18]], paste0(my_path, "Ucumari.xlsx"), row.names = FALSE, sheetName = "dog", append = TRUE)                       # Append to second sheet
write.xlsx2(ucumari_efort[[1]], paste0(my_path, "Ucumari.xlsx"), row.names = FALSE, sheetName = "effort", append = TRUE)                       # Append to third sheet
write.xlsx2(as.data.frame(ucumari), paste0(my_path, "Ucumari.xlsx"), row.names = TRUE, sheetName = "camera", append = TRUE)                       # Append to third sheet


# number of observetions per sp, collapsed to 7 days
# lapply(ylist, sum, na.rm = TRUE)


# leopardus tigrinus 7
# canis 18

############## Pitalito

pitalito_DetHist_list <- lapply(unique(Full_data_pitalito$scientificName), FUN = function(x) {
  detectionHistory(
    recordTable         = Full_data_pitalito, # tabla de registros
    camOp                = pitalito_camop, # Matriz de operación de cámaras
    stationCol           = "camid",
    speciesCol           = "scientificName",
    recordDateTimeCol    = "eventDateTime",
    recordDateTimeFormat  = "%Y-%m-%d %H:%M:%S",
    species              = x,     # la función reemplaza x por cada una de las especies
    occasionLength       = 6, # Colapso de las historias a 10 ías
    day1                 = "survey", # "survey" a specific date, "station", #inicie en la fecha de cada survey
    datesAsOccasionNames = TRUE,
    includeEffort        = TRUE,
    scaleEffort          = FALSE,
    #unmarkedMultFrameInput=TRUE
    timeZone             = "America/Bogota" 
    )
  }
)

# names
names(pitalito_DetHist_list) <- unique(Full_data_pitalito$scientificName)
# Finalmente creamos una lista nueva donde estén solo las historias de detección
pitalito_ylist <- lapply(pitalito_DetHist_list, FUN = function(x) x$detection_history)
# otra lista con effort scaled
pitalito_efort <- lapply(pitalito_DetHist_list, FUN = function(x) x$effort)


# perro 41
# tigrinus 5

write.xlsx2(pitalito_ylist[[5]], paste0(my_path, "Pitalito.xlsx"), row.names = FALSE, sheetName = "tigrinus")                                      # Write xlsx file
write.xlsx2(pitalito_ylist[[41]], paste0(my_path, "Pitalito.xlsx"), row.names = FALSE, sheetName = "dog", append = TRUE)                       # Append to second sheet
write.xlsx2(pitalito_efort[[1]], paste0(my_path, "Pitalito.xlsx"), row.names = FALSE, sheetName = "effort", append = TRUE)                       # Append to third sheet
write.xlsx2(as.data.frame(pitalito), paste0(my_path, "Pitalito.xlsx"), row.names = TRUE, sheetName = "camera", append = TRUE)                       # Append to third sheet




############## La Fe

# lafe_data <- lafe_data |> 
#   filter(camid != "Palmas_Ladera") |> 
#   filter(camid != "Abuel_Ladera") 

lafe_DetHist_list <- lapply(unique(lafe_data$scientificName), FUN = function(x) {
  detectionHistory(
    recordTable         = lafe_data, # tabla de registros
    camOp                = lafe_camop, # Matriz de operación de cámaras
    stationCol           = "camid",
    speciesCol           = "scientificName",
    recordDateTimeCol    = "eventDateTime",
    recordDateTimeFormat  = "%Y-%m-%d %H:%M:%S",
    species              = x,     # la función reemplaza x por cada una de las especies
    occasionLength       = 7, # Colapso de las historias a 10 días
    day1                 = "survey", # "survey" a specific date, "station", #inicie en la fecha de cada survey
    datesAsOccasionNames = TRUE,
    includeEffort        = TRUE,
    scaleEffort          = FALSE,
    #unmarkedMultFrameInput=TRUE
    timeZone             = "America/Bogota" 
    )
  }
)

# names
names(lafe_DetHist_list) <- unique(lafe_data$scientificName)
# Finalmente creamos una lista nueva donde estén solo las historias de detección
lafe_ylist <- lapply(lafe_DetHist_list, FUN = function(x) x$detection_history)
# otra lista con effort scaled
lafe_efort <- lapply(lafe_DetHist_list, FUN = function(x) x$effort)


# perro 2
# tigrinus 3

write.xlsx2(lafe_ylist[[3]], paste0(my_path, "LaFe.xlsx"), row.names = FALSE, sheetName = "tigrinus")                                      # Write xlsx file
write.xlsx2(lafe_ylist[[2]], paste0(my_path, "LaFe.xlsx"), row.names = FALSE, sheetName = "dog", append = TRUE)                       # Append to second sheet
write.xlsx2(lafe_efort[[1]], paste0(my_path, "LaFe.xlsx"), row.names = FALSE, sheetName = "effort", append = TRUE)                       # Append to third sheet
write.xlsx2(as.data.frame(lafe), paste0(my_path, "LaFe.xlsx"), row.names = TRUE, sheetName = "camera", append = TRUE)                       # Append to third sheet



############## Rio Grande

# lafe_data <- lafe_data |> 
#   filter(camid != "Palmas_Ladera") |> 
#   filter(camid != "Abuel_Ladera") 

riogrande_DetHist_list <- lapply(unique(riogrande_data$scientificName), FUN = function(x) {
  detectionHistory(
    recordTable         = riogrande_data, # tabla de registros
    camOp                = riogrande_camop, # Matriz de operación de cámaras
    stationCol           = "camid",
    speciesCol           = "scientificName",
    recordDateTimeCol    = "eventDateTime",
    recordDateTimeFormat  = "%Y-%m-%d %H:%M:%S",
    species              = x,     # la función reemplaza x por cada una de las especies
    occasionLength       = 8, # Colapso de las historias a 10 días
    day1                 = "survey", # "survey" a specific date, "station", #inicie en la fecha de cada survey
    datesAsOccasionNames = TRUE,
    includeEffort        = TRUE,
    scaleEffort          = FALSE,
    #unmarkedMultFrameInput=TRUE
    timeZone             = "America/Bogota" 
    )
  }
)

# names
names(riogrande_DetHist_list) <- unique(riogrande_data$scientificName)
# Finalmente creamos una lista nueva donde estén solo las historias de detección
riogrande_ylist <- lapply(riogrande_DetHist_list, FUN = function(x) x$detection_history)
# otra lista con effort scaled
riogrande_efort <- lapply(riogrande_DetHist_list, FUN = function(x) x$effort)


# perro 2
# tigrinus 3

write.xlsx2(riogrande_ylist[[3]], paste0(my_path, "RioGrande.xlsx"), row.names = FALSE, sheetName = "tigrinus")                                      # Write xlsx file
write.xlsx2(riogrande_ylist[[2]], paste0(my_path, "RioGrande.xlsx"), row.names = FALSE, sheetName = "dog", append = TRUE)                       # Append to second sheet
write.xlsx2(riogrande_efort[[1]], paste0(my_path, "RioGrande.xlsx"), row.names = FALSE, sheetName = "effort", append = TRUE)                       # Append to third sheet
write.xlsx2(as.data.frame(riogrande), paste0(my_path, "RioGrande.xlsx"), row.names = TRUE, sheetName = "camera", append = TRUE)                       # Append to third sheet



############## Ituango

# lafe_data <- lafe_data |> 
#   filter(camid != "Palmas_Ladera") |> 
#   filter(camid != "Abuel_Ladera") 

ituango_DetHist_list <- lapply(unique(ituango_data$scientificName), FUN = function(x) {
  detectionHistory(
    recordTable         = ituango_data, # tabla de registros
    camOp                = ituango_camop, # Matriz de operación de cámaras
    stationCol           = "camid",
    speciesCol           = "scientificName",
    recordDateTimeCol    = "eventDateTime",
    recordDateTimeFormat  = "%Y-%m-%d %H:%M:%S",
    species              = x,     # la función reemplaza x por cada una de las especies
    occasionLength       = 13, # Colapso de las historias a 10 días
     day1                 = "survey", # "survey" a specific date, "station", #inicie en la fecha de cada survey
    datesAsOccasionNames = TRUE,
    includeEffort        = TRUE,
    scaleEffort          = FALSE,
    #unmarkedMultFrameInput=TRUE
    timeZone             = "America/Bogota" 
    )
  }
)

# names
names(ituango_DetHist_list) <- unique(ituango_data$scientificName)
# Finalmente creamos una lista nueva donde estén solo las historias de detección
ituango_ylist <- lapply(ituango_DetHist_list, FUN = function(x) x$detection_history)
# otra lista con effort scaled
ituango_efort <- lapply(ituango_DetHist_list, FUN = function(x) x$effort)


# perro 5
# tigrinus 26

write.xlsx2(ituango_ylist[[26]], paste0(my_path, "Ituango.xlsx"), row.names = FALSE, sheetName = "tigrinus")                                      # Write xlsx file
write.xlsx2(ituango_ylist[[5]], paste0(my_path, "Ituango.xlsx"), row.names = FALSE, sheetName = "dog", append = TRUE)                       # Append to second sheet
write.xlsx2(ituango_efort[[1]], paste0(my_path, "Ituango.xlsx"), row.names = FALSE, sheetName = "effort", append = TRUE)                       # Append to third sheet
write.xlsx2(as.data.frame(ituango), paste0(my_path, "Ituango.xlsx"), row.names = TRUE, sheetName = "camera", append = TRUE)                       # Append to third sheet






chingaza_DetHist_list <- lapply(unique(chingaza_data$scientificName), FUN = function(x) {
  detectionHistory(
    recordTable         = chingaza_data, # tabla de registros
    camOp                = chingaza_camop, # Matriz de operación de cámaras
    stationCol           = "camid",
    speciesCol           = "scientificName",
    recordDateTimeCol    = "eventDateTime",
    recordDateTimeFormat  = "%Y-%m-%d %H:%M:%S",
    species              = x,     # la función reemplaza x por cada una de las especies
    occasionLength       = 13, # Colapso de las historias a 10 días
    day1                 = "survey", # "survey" a specific date, "station", #inicie en la fecha de cada survey
    datesAsOccasionNames = TRUE,
    includeEffort        = TRUE,
    scaleEffort          = FALSE,
    #unmarkedMultFrameInput=TRUE
    timeZone             = "America/Bogota" 
    )
  }
)

# names
names(chingaza_DetHist_list) <- unique(chingaza_data$scientificName)
# Finalmente creamos una lista nueva donde estén solo las historias de detección
chingaza_ylist <- lapply(chingaza_DetHist_list, FUN = function(x) x$detection_history)
# otra lista con effort scaled
chingaza_efort <- lapply(chingaza_DetHist_list, FUN = function(x) x$effort)


# perro 1
# tigrinus 2


write.xlsx2(chingaza_ylist[[2]], paste0(my_path, "Chingaza.xlsx"), row.names = FALSE, sheetName = "tigrinus")                                      # Write xlsx file
write.xlsx2(chingaza_ylist[[1]], paste0(my_path, "Chingaza.xlsx"), row.names = FALSE, sheetName = "dog", append = TRUE)                       # Append to second sheet
write.xlsx2(chingaza_efort[[1]], paste0(my_path, "Chingaza.xlsx"), row.names = FALSE, sheetName = "effort", append = TRUE)                       # Append to third sheet
write.xlsx2(as.data.frame(chingaza), paste0(my_path, "Chingaza.xlsx"), row.names = TRUE, sheetName = "camera", append = TRUE)                       # Append to third sheet




```

## cocha data

```{r}


# fix dates

Full_data_cocha1$start_date <- as.Date(Full_data_cocha1$"camera_trap_start_date", "%Y-%m-%d")
Full_data_cocha1$end_date <- as.Date(Full_data_cocha1$"camera_trap_end_date", "%Y-%m-%d")
Full_data_cocha1$eventDateTime <- Full_data_cocha1$Date_Time #, "%Y-%m-%d")
# remove NA in datetime

# Full_data_cocha1$eventDateTime <- ymd_hms(paste(Full_data_cocha1$"Photo_Date", Full_data_cocha1$"Photo time", sep=" "))
# rename camera id
Full_data_cocha1$camid <- Full_data_cocha1$`camera_trap`



# filter 2021 and make uniques
cocha1_CToperation  <- Full_data_cocha1 |> dplyr::group_by(camid) |> #(array_locID) |> 
                           mutate(minStart=start_date, maxEnd=end_date) |> distinct(Longitude, Latitude, minStart, maxEnd) |> dplyr::ungroup()
# remove one duplicated
# View(CToperation)
# CToperation <- CToperation[-15,]


# Generamos la matríz de operación de las cámaras

cocha1_camop <- cameraOperation(CTtable= cocha1_CToperation, # Tabla de operación
                         stationCol= "camid", # Columna que define la estación
                         setupCol= "minStart", #Columna fecha de colocación
                         retrievalCol= "maxEnd", #Columna fecha de retiro
                         #hasProblems= T, # Hubo fallos de cámaras
                         dateFormat= "%Y-%m-%d")#, #, # Formato de las fechas
                         #cameraCol="Camera_Id")
                         # sessionCol= "Year")



# Generar las historias de detección ---------------------------------------
## remove problem species

Full_data_cocha1$scientificName <- Full_data_cocha1$`Genus Species`

#### remove setups
cocha1_ind <- which(is.na((Full_data_cocha1$scientificName)))
Full_data_cocha1 <- Full_data_cocha1[-cocha1_ind,]

# ind <- which(Ecu_full$scientificName=="Set up")
# Ecu_full <- Ecu_full[-ind,]
# 
# ind <- which(Ecu_full$scientificName=="Blank")
# Ecu_full <- Ecu_full[-ind,]
# 
# ind <- which(Ecu_full$scientificName=="Unidentifiable")
# Ecu_full <- Ecu_full[-ind,]


############### cocha1

cocha1_DetHist_list <- lapply(unique(Full_data_cocha1$scientificName), FUN = function(x) {
  detectionHistory(
    recordTable         = Full_data_cocha1, # abla de registros
    camOp                = cocha1_camop, # Matriz de operación de cámaras
    stationCol           = "camid",
    speciesCol           = "scientificName",
    recordDateTimeCol    = "eventDateTime",
    recordDateTimeFormat  = "%Y-%m-%d %H:%M:%S",
    species              = x,     # la función reemplaza x por cada una de las especies
    occasionLength       = 8, # Colapso de las historias a 10 ías
    day1                 = "survey", # "survey" a specific date, "station", #inicie en la fecha de cada survey
    datesAsOccasionNames = TRUE,
    includeEffort        = TRUE,
    scaleEffort          = FALSE,
    #unmarkedMultFrameInput=TRUE
    timeZone             = "America/Bogota" 
    )
  }
)

# names
names(cocha1_DetHist_list) <- unique(Full_data_cocha1$scientificName)
# Finalmente creamos una lista nueva donde estén solo las historias de detección
cocha1_ylist <- lapply(cocha1_DetHist_list, FUN = function(x) x$detection_history)
# otra lista con effort scaled
cocha1_efort <- lapply(cocha1_DetHist_list, FUN = function(x) x$effort)

# number of observetions per sp, collapsed to 7 days
# lapply(ylist, sum, na.rm = TRUE)


# leopardus tigrinus 9
# canis 1


write.xlsx2(cocha1_ylist[[9]], paste0(my_path, "Cocha1.xlsx"), row.names = FALSE, sheetName = "tigrinus")                                      # Write xlsx file
write.xlsx2(cocha1_ylist[[1]], paste0(my_path, "Cocha1.xlsx"), row.names = FALSE, sheetName = "dog", append = TRUE)                       # Append to second sheet
write.xlsx2(cocha1_efort[[1]], paste0(my_path, "Cocha1.xlsx"), row.names = FALSE, sheetName = "effort", append = TRUE)                       # Append to third sheet
write.xlsx2(as.data.frame(cocha1), paste0(my_path, "Cocha1.xlsx"), row.names = TRUE, sheetName = "camera", append = TRUE)                       # Append to third sheet



############################
#### Cocha 2
############################


# fix dates

Full_data_cocha2$start_date <- as.Date(Full_data_cocha2$"camera_trap_start_date", "%Y-%m-%d")
Full_data_cocha2$end_date <- as.Date(Full_data_cocha2$"camera_trap_end_date", "%Y-%m-%d")

Full_data_cocha2$eventDateTime <- ymd_hms(paste(Full_data_cocha2$"Photo Date", Full_data_cocha2$"Photo time", sep=" "))

# remove NA in datetime

# Full_data_cocha1$eventDateTime <- ymd_hms(paste(Full_data_cocha1$"Photo_Date", Full_data_cocha1$"Photo time", sep=" "))
# rename camera id
Full_data_cocha2$camid <- Full_data_cocha2$`Camera Trap Name`



# filter 2021 and make uniques
cocha2_CToperation  <- Full_data_cocha2 |> dplyr::group_by(camid) |> #(array_locID) |> 
                           mutate(minStart=start_date, maxEnd=end_date) |> distinct(Longitude, Latitude, minStart, maxEnd) |> dplyr::ungroup()
# remove one duplicated
# View(CToperation)
# CToperation <- CToperation[-15,]


# Generamos la matríz de operación de las cámaras

cocha2_camop <- cameraOperation(CTtable= cocha2_CToperation, # Tabla de operación
                         stationCol= "camid", # Columna que define la estación
                         setupCol= "minStart", #Columna fecha de colocación
                         retrievalCol= "maxEnd", #Columna fecha de retiro
                         #hasProblems= T, # Hubo fallos de cámaras
                         dateFormat= "%Y-%m-%d")#, #, # Formato de las fechas
                         #cameraCol="Camera_Id")
                         # sessionCol= "Year")



# Generar las historias de detección ---------------------------------------
## remove problem species

Full_data_cocha2$scientificName <- Full_data_cocha2$`Genus Species`

#### remove setups
cocha2_ind <- which(is.na((Full_data_cocha2$scientificName)))
Full_data_cocha2 <- Full_data_cocha2[-cocha2_ind,]

# ind <- which(Ecu_full$scientificName=="Set up")
# Ecu_full <- Ecu_full[-ind,]

############### cocha2

cocha2_DetHist_list <- lapply(unique(Full_data_cocha2$scientificName), FUN = function(x) {
  detectionHistory(
    recordTable         = Full_data_cocha2, # abla de registros
    camOp                = cocha2_camop, # Matriz de operación de cámaras
    stationCol           = "camid",
    speciesCol           = "scientificName",
    recordDateTimeCol    = "eventDateTime",
    recordDateTimeFormat  = "%Y-%m-%d %H:%M:%S",
    species              = x,     # la función reemplaza x por cada una de las especies
    occasionLength       = 13, # Colapso de las historias a 10 ías
    day1                 = "survey", # "survey" a specific date, "station", #inicie en la fecha de cada survey
    datesAsOccasionNames = TRUE,
    includeEffort        = TRUE,
    scaleEffort          = FALSE,
    #unmarkedMultFrameInput=TRUE
    timeZone             = "America/Bogota" 
    )
  }
)


# names
names(cocha2_DetHist_list) <- unique(Full_data_cocha2$scientificName)
# Finalmente creamos una lista nueva donde estén solo las historias de detección
cocha2_ylist <- lapply(cocha2_DetHist_list, FUN = function(x) x$detection_history)
# otra lista con effort scaled
cocha2_efort <- lapply(cocha2_DetHist_list, FUN = function(x) x$effort)

# number of observetions per sp, collapsed to 7 days
# lapply(ylist, sum, na.rm = TRUE)


# leopardus tigrinus 29
# canis NA


write.xlsx2(cocha2_ylist[[29]], paste0(my_path, "Cocha2.xlsx"), row.names = FALSE, sheetName = "tigrinus")                                      # Write xlsx file
write.xlsx2(cocha2_ylist[[1]], paste0(my_path, "Cocha2.xlsx"), row.names = FALSE, sheetName = "dog", append = TRUE)                       # Append to second sheet
write.xlsx2(cocha2_efort[[1]], paste0(my_path, "Cocha2.xlsx"), row.names = FALSE, sheetName = "effort", append = TRUE)                       # Append to third sheet
write.xlsx2(as.data.frame(cocha2), paste0(my_path, "Cocha2.xlsx"), row.names = TRUE, sheetName = "camera", append = TRUE)                       # Append to third sheet




# Pitalito
# perro 41
# tigrinus 5



```

## Data assembly

```{r}
#### Cocha 2
tigrinus_cocha2 <- cocha2_ylist[[29]] |> as.data.frame()
effort_cocha2 <- cocha2_efort[[29]] |>  as.data.frame() |> 
  mutate(across(everything(), ~replace_na(., 0))) # replace NA to 0 

# perro no hay en cocha2
my_vector <- cocha2_ylist[[29]]
perros_cocha2 <- ifelse(my_vector == 1, 0, my_vector)  |> as.data.frame() # convert 1 to 0

#### Cocha 1
tigrinus_cocha1 <- cocha1_ylist[[9]] |> as.data.frame()
effort_cocha1 <- cocha1_efort[[9]] |> as.data.frame() |> 
  mutate(across(everything(), ~replace_na(., 0))) # replace NA to 0 
# si hay perro en cocha2
# my_vector <- tigrinus_cocha2
perros_cocha1 <- cocha1_ylist[[1]] |> as.data.frame()# ifelse(my_vector == 1, 0, my_vector) # convert 1 to 0

#### LaFe
tigrinus_lafe <- lafe_ylist[[3]] |> as.data.frame()
effort_lafe <- lafe_efort[[3]] |> as.data.frame()|> 
  mutate(across(everything(), ~replace_na(., 0))) # replace NA to 0 
# my_vector <- tigrinus_cocha2
perros_lafe <- lafe_ylist[[2]] |> as.data.frame() # ifelse(my_vector == 1, 0, my_vector) # convert 1 to 0

# perro 2
# tigrinus 3


#### Rio Grande
tigrinus_riogrande <- riogrande_ylist[[11]] |> as.data.frame()
effort_riogrande <- riogrande_efort[[11]] |> as.data.frame()|> 
  mutate(across(everything(), ~replace_na(., 0))) # replace NA to 0 
# my_vector <- tigrinus_cocha2
perros_riogrande <- riogrande_ylist[[1]] |> as.data.frame() # ifelse(my_vector == 1, 0, my_vector) # convert 1 to 0

# perro 1
# tigrinus 11


#### Ituango
tigrinus_ituango <- ituango_ylist[[26]] |> as.data.frame()
effort_ituango <- ituango_efort[[26]] |> as.data.frame()|> 
  mutate(across(everything(), ~replace_na(., 0))) # replace NA to 0 
# my_vector <- tigrinus_cocha2
perros_ituango <- ituango_ylist[[5]] |> as.data.frame() # ifelse(my_vector == 1, 0, my_vector) # convert 1 to 0

# perro 5
# tigrinus 26




#### Chingaza
tigrinus_chingaza <- chingaza_ylist[[2]] |> as.data.frame()
effort_chingaza <- chingaza_efort[[2]] |> as.data.frame()|> 
  mutate(across(everything(), ~replace_na(., 0))) # replace NA to 0 
# my_vector <- tigrinus_cocha2
perros_chingaza <- chingaza_ylist[[1]] |> as.data.frame() # ifelse(my_vector == 1, 0, my_vector) # convert 1 to 0

# perro 1
# tigrinus 2




#### Pitalito

tigrinus_pitalito <- pitalito_ylist[[5]] |> as.data.frame()
effort_pitalito <- pitalito_efort[[5]] |> as.data.frame()|> 
  mutate(across(everything(), ~replace_na(., 0))) # replace NA to 0 
# my_vector <- tigrinus_cocha2
perros_pitalito <- pitalito_ylist[[41]] |> as.data.frame() # ifelse(my_vector == 1, 0, my_vector) # convert 1 to 0

# perro 41
# tigrinus 5

######### Ucumari

tigrinus_ucumari <- ucumari_ylist[[7]] |> as.data.frame()
effort_ucumari <- ucumari_efort[[7]] |> as.data.frame()|>
  mutate(across(everything(), ~replace_na(., 0))) # replace NA to 0 
# my_vector <- tigrinus_cocha2
perros_ucumari <- ucumari_ylist[[18]] |> as.data.frame() # ifelse(my_vector == 1, 0, my_vector) # convert 1 to 0

# leopardus tigrinus 7
# canis 18


################# Saldania
tigrinus_saldania <- tigrinus_Saldania[1:11] |> 
  column_to_rownames(var = "...1") |> 
  as.data.frame() 

perros_saldania <- perro_Saldania[1:11] |> 
  column_to_rownames(var = "...1") |> 
  as.data.frame()

effort_Saldania <- tigrinus_Saldania[15:24] |> 
  # column_to_rownames(var = "...1") |> 
  as.data.frame()
row.names(effort_Saldania) <- row.names(tigrinus_saldania)

names(tigrinus_saldania) <- c("o1","o2","o3","o4","o5", "o6","o7","o8","o9","o10")

names(perros_saldania) <- c("o1","o2","o3","o4","o5", "o6","o7","o8","o9","o10")

names(effort_Saldania) <- c("o1","o2","o3","o4","o5", "o6","o7","o8","o9","o10")


# fix tigrinus_pitalito to complete 11 ocasiones
# tigrinus_pitalito$o11 <- NA
# perros_pitalito$o11 <- NA
# effort_pitalito$o11 <- NA
 
# fix tigrinus_lafe to complete 11 ocasiones
# tigrinus_lafe$o11 <- NA
# perros_lafe$o11 <- NA
# effort_lafe$o11 <- NA

DL_tigrinus <- rbind(tigrinus_ucumari,
                     tigrinus_pitalito,
                     tigrinus_cocha1,
                     tigrinus_lafe,
                     tigrinus_riogrande,
                     tigrinus_ituango,
                     tigrinus_cocha2,
                     tigrinus_chingaza,
                     tigrinus_saldania)
                     # tigrinus_cocha2
                     #)

DL_perros <- rbind(perros_ucumari,
                   perros_pitalito,
                   perros_cocha1,
                   perros_lafe,
                   perros_riogrande,
                   perros_ituango,
                   perros_cocha2,
                   perros_chingaza,
                   perros_saldania)

DL_effort <- rbind(effort_ucumari,
                effort_pitalito,
                effort_cocha1,
                effort_lafe,
                effort_riogrande,
                effort_ituango,
                perros_cocha2,
                perros_chingaza,
                effort_Saldania)

# add colname to later extract  covs
DL_tigrinus$camera_trap <- row.names(DL_tigrinus)
DL_perros$camera_trap <- row.names(DL_perros)
DL_effort$camera_trap <- row.names(DL_effort)

# Letf join con puntos
DL_tigrinus_p <- left_join(DL_tigrinus, puntos)
DL_perros_p <- left_join(DL_perros, puntos)
# DL_tigrinus_p <- left_join(DL_tigrinus, puntos)

########## add spatial covs

# make sf and add projection
puntos_tigrinus_sf <- DL_tigrinus_p |> st_as_sf(coords = 
                              c("Longitude", "Latitude"), 
           crs = "EPSG:4326")

# Extract coordinates and drop geometry
# coordinates <- st_coordinates(puntos_sf)
# data_no_geometry <- st_drop_geometry(sf_data)


# get elevation points... slow!
site_covs_ucu <- get_elev_point(puntos_tigrinus_sf[1:61,], 
                                src = "aws", z = 12)
site_covs_pit <- get_elev_point(puntos_tigrinus_sf[62:122,], 
                                src = "aws", z = 12)
site_covs_coc1 <- get_elev_point(puntos_tigrinus_sf[123:165,], 
                                src = "aws", z = 12)
# site_covs_coc2 <- get_elev_point(puntos_tigrinus_sf[166:218,], 
#                                 src = "aws", z = 12)
site_covs_lafe <- get_elev_point(puntos_tigrinus_sf[166:192,], 
                                src = "aws", z = 12)
site_covs_riogrande <- get_elev_point(puntos_tigrinus_sf[193:224,], 
                                src = "aws", z = 12)
site_covs_ituango <- get_elev_point(puntos_tigrinus_sf[225:278,], 
                                src = "aws", z = 12)
site_covs_coc2 <- get_elev_point(puntos_tigrinus_sf[279:331,], 
                                src = "aws", z = 12)
site_covs_ching <- get_elev_point(puntos_tigrinus_sf[332:356,], 
                                src = "aws", z = 12)
site_covs_salda <- get_elev_point(puntos_tigrinus_sf[357:387,], 
                                src = "aws", z = 12)

# combine points in one sf object
site_covs <- rbind(site_covs_ucu, 
                   site_covs_pit, 
                   site_covs_coc1,
                   site_covs_lafe,
                   site_covs_riogrande,
                   site_covs_ituango,
                   site_covs_coc2,
                   site_covs_ching,
                   site_covs_salda)#, 
#                   site_covs_coc2)
                   #z =1-14
# bb <-  st_as_sfc(st_bbox(elevation_17)) # make bounding box 




############## make distance map using SF
# Convert points to sp spatialpointdatafram
# casas_points <- as(casas_sf, "Spatial")
# Projection

casas_points_utm <- st_transform(casas_sf, CRS('+init=epsg:21818'))
# convert sf to ppp

nc_spatvect <- vect(casas_points_utm)
c_spatvect  <- vect(casas_points_utm)
casa_dist_rast <- distance(rast(nc_spatvect, resolution = 100), c_spatvect) #|> mask(nc_spatvect) # < resolution + detail

# Extrae distancia casas
site_covs$dist_casa <- raster::extract(casa_dist_rast, site_covs)[,2] # also works


# rename elevation
site_covs <- site_covs |> rename(elev = elevation)

### overwrite Casas Chingaza
chi_cov_occ <- read_csv("D:/BoxFiles/Box Sync/CodigoR/tigrinus2/data/chi_cov_occ.csv") |> 
  rename(dist_casa=dist_casas) |> 
  rename(camera_trap=Camara) |> select(c(dist_casa,camera_trap))
# View(chi_cov_occ)

site_covs2 <- site_covs |> 
  filter (region=="Chingaza NP") |> 
  left_join(chi_cov_occ, by = "camera_trap")
# mapview(casa_dist_rast) + mapview(site_covs[,"camera_trap"])

site_covs3 <- site_covs |> 
  filter (region=="Ituango") |> 
  left_join(ituango_dis_casa, by = "camera_trap")
# mapview(casa_dist_rast) + mapview(site_covs[,"camera_trap"])


### dirty add for Chingaza e Ituango dist_casa
site_covs$dist_casa[332:356] <- site_covs2$dist_casa.y #chingaza
site_covs$dist_casa[225:278] <- replace_na(site_covs3$dist_casa.y, 610) # Ituango



##############
# Canopy High
##############

ucu_point <- puntos_tigrinus_sf |> filter (region=="Ucumari RP")

ucu_pol <- st_as_sf( # make sf
             st_buffer( # bufer 1 k
               st_as_sfc(# make sfc
                 st_bbox(ucu_point) # box around point
                 ), 
              dist = 500)# distance bufer 500m
             )


library(terra)
library(tidyterra)
library(aws.s3)
library(forestdata)
## Download the data
# very high resolution canopy height maps 10m
ucumari_eth <- fd_canopy_height(
  x     = ucu_pol,
  model = "eth",
  crop  = TRUE
  #mask  = TRUE
)


## Download the GEDI forest data... takes time!
#  0-60 Forest canopy height, meters
# 30 m/pixel spatial resolution
# https://doi.org/10.1016/j.rse.2020.112165
ucumari_forest <- fd_forest_glad( #
  x=ucu_pol,
  #lon = -75.53673, lat = 4.748044,
    year = 2020,
  model  = "landcover",
  # model = "eth",
  crop  = TRUE,
  mask  = TRUE
)



## Print raster
#print(ucumari_forest)





```

## Map

### General map

```{r}
#
# fom rnaturalearth
col <- ne_countries(country = "Colombia", scale = "medium")

# general
tm_shape(col)+
tm_shape(puntos_tigrinus_sf, # add bb
         bbox = tmaptools::bb(puntos_tigrinus_sf, ext = 1.5)) +
    tm_basemap("OpenStreetMap") + # c(StreetMap = "OpenStreetMap", TopoMap = "OpenTopoMap")) +# ("Esri.WorldImagery") + # usa basemap
   tm_symbols(shape = 21, col = "red", fill = "blue",size =0.4) #+ #punto negro
  # tm_facets(by = "region", ncol = 3)

```

### Zoom to regions

```{r}
#| column: body-outset
#| 

#####| column: screen-inset-shaded

### detallado
tm_shape(puntos_tigrinus_sf) +
    tm_basemap("OpenTopoMap") +# ("Esri.WorldImagery") + # usa basemap
   tm_symbols(shape = 21, col = "red", fill = "blue",size =0.4) + #punto negro
  tm_facets(by = "region", ncol = 3)

```

## MODELOS DE CO-OCURRENCIA

```{r}
#| output: true

detformulas <- c( "~1", "~1")#, "~1")
#stateformulas <- c('~elev','~elev', '~elev', "~1", "~1", "~1", "~0")# 3 sp
stateformulas <- c('~elev','~elev', "~1") #"~0"
stateformulas_2 <- c('~elev + I(elev^2)','~elev', "~1") #"~0"

y <- list(as.matrix(DL_tigrinus_p[,1:10]),# truncate 10
          as.matrix(DL_perros_p[,1:10])
          )#[,1:5]))# , tigrinus y perros)
names(y) <- c("tigrinus", "perros")#, "ocelote")


# obs_covs <-as.data.frame(scale(cams_ucu_sf$dist_casa))
# names(obs_covs) <- "dist_casa"

site_covs_ucu <- data.frame(site_covs[,c('elev','dist_casa')])[,1:2]
site_covs_ucu <-as.data.frame(apply(site_covs_ucu,2,scale)) # notice scale here
names(site_covs_ucu) <- c("elev", "dist_casa")


#### effort
obs_covs <- list(
  efort_tig=as.data.frame(DL_effort[,1:10]), # truncate to 10
  effort_perro=as.data.frame(DL_effort[,1:10])
)


library(unmarked)

umf <-  unmarkedFrameOccuMulti(y = y, 
                              siteCovs = site_covs_ucu,
                              obsCovs  = obs_covs)#NULL)
plot(umf)
#umf

# occFormulas Length should match number/order of columns in fDesign
umf@fDesign

#########################
null_det <- c("~1", "~1")
null_occu <- c("~1", "~1")#, "~0")
null <- occuMulti(null_det, null_occu, 
                  umf,
                  method="BFGS", 
                  se=TRUE, 
                  engine=c("C"), 
                  silent=TRUE,
                  maxOrder=1, 
                  penalty =2,#0.5 * sum(paramvals^2)
                  boot=500
                  )

null

########################
fit1 <- occuMulti(detformulas, stateformulas, umf,    
        method="BFGS", se=TRUE, engine=c("C"), silent=TRUE,
        maxOrder=2, 
                  penalty =2,#0.5 * sum(paramvals^2)
                  boot=500)

fit1



# update model
# occFormulas2 <- c('~dist_casa', '~dist_casa', '~dist_casa', "~1", "~1", "~1", "~0")
occFormulas2 <- c('~dist_casa', '~dist_casa', "~1")
fit2 <- update(fit1, stateformulas=occFormulas2)
fit2

detFormulas_eff <- c('~efort_tig', '~effort_perro')
fit3 <- update(fit1, detformulas=detFormulas_eff)
fit3

fit4 <- update(fit2, detformulas=detFormulas_eff)


fit1_2 <- occuMulti(detFormulas_eff, stateformulas_2, umf,    
        method="BFGS", se=TRUE, engine=c("C"), silent=TRUE,
        maxOrder=2, 
                  penalty =2,#0.5 * sum(paramvals^2)
                  boot=500)



# null_2order <- optimizePenalty(null, 
#                                 stateformulas = c("~1", "~1", "~0"),
#                                 penalties = 2, # c(0, 2^seq(-4, 4))
#                                 maxOrder=2,
#                                 k = 5, boot = 250)
# summary(null_2order)


```

### Model Selection

```{r}
#| output: asis
#| 
#List of fitted models
fmList <- fitList(#elev = fit1, 
                  #distCasa = fit2, 
                  elev_effort = fit3,
                  distCasa_effort = fit4,
                  # null_2order = null_2order,
                  null = null,
                  elev2_eff = fit1_2)


#Model selection
models <- modSel(fmList)


models_toExport <- as(models, "data.frame") |> select(c(model,nPars,AIC,delta,AICwt,cumltvWt)) 


# print table
DT::datatable(models_toExport) |>  formatRound(c('AIC',"delta", "AICwt", "cumltvWt"), 3)

# coef(fmList)





```

#### Best model fit

```{r}
#############
# Model fit #
#############

bt <- parboot(fit1_2, nsim=100) # takes time best model
plot(bt)
```

### plot predicted marginal occupancy

look at occupancy for species individually.

```{r}
#Plot predicted marginal occupancy as a function of disturbance
r <- range(site_covs$elev)
x1 <- seq(r[1],r[2],length.out=100)
x_scale <- (x1-mean(site_covs$elev))/sd(site_covs$elev)

r2 <- range(site_covs$dist_casa)
x2 <- seq(r2[1],r2[2],length.out=100)
x2_scale <- (x2-mean(site_covs$dist_casa))/sd(site_covs$dist_casa)

nd <- matrix(NA, 100, 2)
nd <- data.frame(elev=x_scale, dist_casa= x2_scale)



tigrinus_pred <- predict(fit1_2, "state", species="tigrinus", newdata=nd)
tigrinus_pred$Species <- "tigrinus"

perros_pred <- predict(fit1_2, "state", species="perros", newdata=nd)
perros_pred$Species <- "perros"

# ocelote_pred <- predict(fit2, "state", species="ocelote", newdata=nd)
# ocelote_pred$Species <- "ocelote"






################### point plot
############## Null model 
tigrinus_pred_null <- predict(fit3, "state", species="tigrinus")
tigrinus_pred_null$Species <- "tigrinus"

perros_pred_null <- predict(fit3, "state", species="perros")
perros_pred_null$Species <- "perros"

all_marginal <- rbind(tigrinus_pred_null[1,], perros_pred_null[1,])
all_marginal$Species <- c("tigrinus", "perros")

#plot
plot(1:2, all_marginal$Predicted, ylim=c(0,0.9), 
     xlim=c(0.5,2.5), pch=19, cex=1.5, xaxt='n', 
     xlab="", ylab="Marginal occupancy and 95% CI")
axis(1, at=1:2, labels=all_marginal$Species)

# CIs
top <- 0.1
for (i in 1:2){
  segments(i, all_marginal$lower[i], i, all_marginal$upper[i])
  segments(i-top, all_marginal$lower[i], i+top)
  segments(i-top, all_marginal$upper[i], i+top)
}



#################################

plot_dat <- rbind(tigrinus_pred, perros_pred)#, ocelote_pred)

ggplot(data=plot_dat, aes(x=rep(x1,2), y=Predicted)) + # change to 3 sp and x2 to distance
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Species), alpha=0.3) +
  geom_line(aes(col=Species)) +
  labs(x="elevation", y="Marginal occupancy") +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        axis.text=element_text(size=12), axis.title=element_text(size=14),
        legend.text=element_text(size=12), legend.title=element_text(size=14))



```

### plot predicted co-occurrence occupancy

```{r}
#Plot predicted marginal occupancy as a function of disturbance
r <- range(site_covs$elev)
x1 <- seq(r[1],r[2],length.out=100)
x_scale <- (x1-mean(site_covs$elev))/sd(site_covs$elev)

r2 <- range(site_covs$dist_casa)
x2 <- seq(r2[1],r2[2],length.out=100)
x2_scale <- (x2-mean(site_covs$dist_casa))/sd(site_covs$dist_casa)

nd <- matrix(NA, 100, 2)
nd <- data.frame(elev=x_scale, dist_casa= x2_scale)


tigrinus_perros_pred <- predict(fit1_2, "state", 
                         species=c("tigrinus", "perros"), 
                         # cond=c('-perros'), #perro absent
                         newdata=nd)

# tigrinus_pred$Species <- c("tigrinus", "perros")

# perros_pred <- predict(fit1, "state", species="perros", newdata=nd)
# perros_pred$Species <- "perros"

# ocelote_pred <- predict(fit2, "state", species="ocelote", newdata=nd)
# ocelote_pred$Species <- "ocelote"


plot_dat <- tigrinus_perros_pred #rbind(tigrinus_pred, perros_pred)#, ocelote_pred)

ggplot(data=plot_dat, aes(x=rep(x1), y=Predicted)) + # change to 3 sp and x2 to distance
  geom_ribbon(aes(ymin=lower, ymax=upper, fill = "grey50"), alpha=0.3) +
  geom_line(aes(y=Predicted), col="blue") +
  labs(x="elev", y="co-occurence") +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        axis.text=element_text(size=12), axis.title=element_text(size=14),
        legend.text=element_text(size=12), legend.title=element_text(size=14))



```

### plot predicted conditional occupancy

you want to know the probability of occupancy of one species, conditional on the presence of another

```{r}
#Plot predicted marginal occupancy as a function of disturbance
r <- range(site_covs$elev)
x1 <- seq(r[1],r[2],length.out=100)
x_scale <- (x1-mean(site_covs$elev))/sd(site_covs$elev)

r2 <- range(site_covs$dist_casa)
x2 <- seq(r2[1],r2[2],length.out=100)
x2_scale <- (x2-mean(site_covs$dist_casa))/sd(site_covs$dist_casa)

nd <- matrix(NA, 100, 2)
nd <- data.frame(elev=x_scale, dist_casa= x2_scale)

##############
######## point plot
######## null model
##############

tigrinus_con_perro <- predict(fit1_2, type="state", species="tigrinus", cond="perros")

tigrinus_no_perro <- predict(fit1_2, type="state", species="tigrinus", cond="-perros")

cond_data <- rbind(tigrinus_con_perro[1,], tigrinus_no_perro[1,])
cond_data$tigrinus_status <- c("Present","Absent")

plot(1:2, cond_data$Predicted, ylim=c(0.15,0.9), 
     xlim=c(0.5,2.5), pch=19, cex=1.5, xaxt='n', 
     xlab="Perro status", ylab="tigrinus occupancy and 95% CI")
axis(1, at=1:2, labels=cond_data$tigrinus_status)

# CIs
top <- 0.1
for (i in 1:2){
  segments(i, cond_data$lower[i], i, cond_data$upper[i])
  segments(i-top, cond_data$lower[i], i+top)
  segments(i-top, cond_data$upper[i], i+top)
}

##############
# new data conditional
nd_cond <- data.frame(
  #elev = rep(mean(site_covs$elev), 100),
  dist_casa = rep(mean(x2_scale), 100),
  # roads = rep(mean(site_covs$roads), 100),
  elev = seq(min(x_scale), max(x_scale),
                 length.out = 100)
)

##### conditional

tigrinus_dog_0 <- predict(fit1_2, "state", 
                         species = "tigrinus", 
                         cond = '-perros',
                         newdata = nd_cond)

tigrinus_dog_0$Species <- "perro ausente"

tigrinus_dog_1 <- predict(fit1_2, "state", 
                         species = "tigrinus", 
                         cond = 'perros',
                         newdata = nd_cond)

tigrinus_dog_1$Species <- "perro presente"





perros_pred <- predict(fit1, "state", species = "perros", newdata = nd)
perros_pred$Species <- "perros"

# ocelote_pred <- predict(fit2, "state", species="ocelote", newdata=nd)
# ocelote_pred$Species <- "ocelote"

# old plot
plot_dat <- rbind(tigrinus_dog_1, tigrinus_dog_0)#, ocelote_pred)

ggplot(data=plot_dat, aes(x=rep(x1,2), y=Predicted)) + # change to 3 sp and x2 to distance
   geom_ribbon(aes(ymin=lower, ymax=upper, fill=Species), alpha=0.3) +
   geom_line(aes(col=Species)) +
   labs(x="elev", y="tigrinus conditional occupancy") +
   theme_bw() +
   theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
         axis.text=element_text(size=12), axis.title=element_text(size=14),
         legend.text=element_text(size=12), legend.title=element_text(size=14))

# new plot




gg_df_cond1 <- data.frame(
  elev = rep(nd_cond$elev, 2),
  occupancy = c(tigrinus_dog_1$Predicted,
                tigrinus_dog_0$Predicted),
  low = c(tigrinus_dog_1$lower,
          tigrinus_dog_0$lower),
  high = c(tigrinus_dog_1$upper,
           tigrinus_dog_0$upper),
  conditional = rep(c('Dog present', 'Dog absent'),
                    each = 100)
)


cond_fig1 <- ggplot(gg_df_cond1, aes(x = elev, y = occupancy,
                                   group = conditional)) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = conditional),  alpha=0.5) +
  geom_line() +
  ylab('Conditional L. tigrinus\noccupancy probability') +
  xlab('elev') +
  labs(fill = 'Dog state') +
  theme(text = element_text(size = 15),
        #legend.position = c(0.75, 0.85)
        )

cond_fig1


```

## Package Citation

```{r}
#| code-fold: true
pkgs <- cite_packages(output = "paragraph", pkgs="Session", out.dir = ".")
# knitr::kable(pkgs)
pkgs
```

## Sesion info

::: {.callout-note collapse="true"}
```{r}
print(sessionInfo(), locale = FALSE)
```
:::
