---
project:
  type: website

website:  
  title: "Occupación tigrinus"
  subtitle: "co-ocurrencia y actividad"
  author: "Diego J. Lizcano & Alejandra Bonilla"
  page-navigation: true
  navbar:
    right:
      - icon: github
        href: https://github.com/
        aria-label: GitHub

execute: 
  echo: true
  eval: true
  warning: false
  error: false
  output: asis
  
format: 
  html: 
    code-fold: true
    code-summary: "codigo R"
    fig-width: 8
    fig-height: 6
---

# cargar paquetes

```{r}
library(knitr)
library(tidyverse) # maneja datos
library(mapview) # mapas facil
library(readxl) #leer datos
library(sf) # vector map
library(geodata) # replace getData de raster para Terra
library(raster) # mapas raster
library(spatstat) # interpola mapa
library(maptools) # to coerce to ppp. note that 'maptools' will be retired by the end of 2023
# library(stars)
library(unmarked) # occu models
library(DT) # html table


source("D:/BoxFiles/Box Sync/CodigoR/tigrinus/R/organizadato.R")

```

# UCUMARI

## cargar datos

```{r}


Full_data_ucu <- read_excel("D:/BoxFiles/Box Sync/CodigoR/tigrinus/data/Full_data_Ucumari_Huila_Cocha1_Cocha2.xlsx", 
    sheet = "ucumari", col_types = c("numeric", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "numeric", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "numeric", 
        "numeric", "numeric", "numeric", 
        "text"))



```

## crerar matrices para unmarked

```{r}
#| column: screen-inset-shaded


casas <- read_csv("D:/BoxFiles/Box Sync/CodigoR/tigrinus/data/casas.csv")
casas_sf <- st_as_sf(casas, coords = c("lon", "lat"), crs = "EPSG:4326")




############# start spatial part
#### make sf object
ucumari_sf <- st_as_sf(Full_data_ucu, coords = c("Longitude", "Latitude"), crs = "EPSG:4326")


# get centroid. 1st bbox, make pol
centroid_ucu <- st_centroid(st_as_sfc(st_bbox(ucumari_sf)))
# get altitude
elev_ucu_full <- elevation_3s(centroid_ucu[[1]][1], centroid_ucu[[1]][2], 
                         path="D:/BoxFiles/Box Sync/CodigoR/tigrinus/raster")

# elev_ucu_full_ras <- elev_ucu_full %>% raster()

# st_bbox(ucumari_sf) # notice order xmin, xmax, ymin, ymax
ext_ucu <- ext(-75.59, -75.47,  4.68,    4.81 )
elev_ucu <- crop(elev_ucu_full, ext_ucu) 

# convert from terra to raster
elev_ucu_ras <-  elev_ucu %>% raster()

# get uniques
cams_ucu <-Full_data_ucu %>% dplyr::select(c("Longitude", "Latitude", "camera_trap")) %>% distinct()
#### make sf object
cams_ucu_sf <- st_as_sf(cams_ucu, coords = c("Longitude", "Latitude"), crs = "EPSG:4326")

# extract values from raster using altitude
cams_ucu_sf$elev <- extract(elev_ucu_ras, cams_ucu_sf)

############## make distance map
# Convert points to sp spatialpointdatafram
casas_points <- as(casas_sf, "Spatial")
# Projection
# Be sure to have rgdal first installed.
casas_points_utm <- spTransform(casas_points, CRS('+init=epsg:32718'))
# convert to ppp
casas_points_ppp <- as(as(casas_points_utm, "SpatialPoints"), "ppp")
# distance
casas_distance <- distmap(casas_points_ppp)

####### extract distance 

casas_distance_ras<- raster(casas_distance) # convert raster
crs(casas_distance_ras) <- '+init=epsg:32718' # add crs

# project raster
casa_distance <- projectRaster(casas_distance_ras, crs = crs(casas_points))

# cams_ucu_sf$dist_casa <- raster::extract(casas_distance_ras, cams_ucu_sf)
cams_ucu_sf$dist_casa <- raster::extract(casa_distance, cams_ucu_sf) # also works


# plot map
mapview(elev_ucu_ras) + mapview(cams_ucu_sf["camera_trap"]) 

############### end spatial part

Full_data_ucu$binomial <- str_c (Full_data_ucu$Genus, "_", Full_data_ucu$Species)

#funcion para crear todas las tablas de datos
all_data <-  f.matrix.creator2 (Full_data_ucu)

# names(all_data) # ver lass especies y en que lista esta cada una
# kable(names(all_data)) # html table
# Tigrinus es lista 8

datatable(
data = as.data.frame(names(all_data)),
caption = "Especies Ucumari",
filter = "top"
)


```

## unmarked

```{r}

# tabla con solo tiginus
tigrinus_ucu <- all_data[[8]]
# cargar paquete
library(unmarked)
# crear objeto umf
umf_tigrinus_ucu <- unmarkedFrameOccu(y=tigrinus_ucu)
# verificar datos en grafica
# plot(umf_tigrinus_ucu)

# tabla con solo perros
perros_ucu <- all_data[[75]]
# crear objeto umf
umf_perros_ucu <- unmarkedFrameOccu(y=perros_ucu)
# verificar datos en grafica
# plot(umf_perros_ucu)

# tabla con solo ocelote
ocelote_ucu <- all_data[[16]]
# crear objeto umf
umf_ocelote_ucu <- unmarkedFrameOccu(y=ocelote_ucu)
# verificar datos en grafica
# plot(umf_ocelote_ucu)

```

## Modelo nulo

```{r}
# modelo nulo tigrinus
fm_tig_ucu <- occu(~1 ~1, umf_tigrinus_ucu)  # fit a model

backTransform(fm_tig_ucu, type="det") #estimado lineal de deteccion
backTransform(fm_tig_ucu, type="state") # estimado linel de ocupacion

# modelo nulo perro
fm_perros_ucu <- occu(~1 ~1, umf_perros_ucu)  # fit a model

backTransform(fm_perros_ucu, type="det") #estimado lineal de deteccion
backTransform(fm_perros_ucu, type="state") # estimado linel de ocupacion

# modelo nulo ocelote
fm_ocelote_ucu <- occu(~1 ~1, umf_ocelote_ucu)  # fit a model

backTransform(fm_ocelote_ucu, type="det") #estimado lineal de deteccion
backTransform(fm_ocelote_ucu, type="state") # estimado linel de ocupacion


```

## MODELOS DE CO-OCURRENCIA

```{r}

detformulas <- c( "~1", "~1")#, "~1")
#stateformulas <- c('~elev','~elev', '~elev', "~1", "~1", "~1", "~0")# 3 sp
stateformulas <- c('~elev','~elev', "~0")
y <- list(tigrinus_ucu, perros_ucu)# , ocelote_ucu)
names(y) <- c("tigrinus", "perros")#, "ocelote")


obs_covs <-as.data.frame(scale(cams_ucu_sf$dist_casa))
names(obs_covs) <- "dist_casa"

site_covs <- data.frame(cams_ucu_sf[,c('elev','dist_casa')])[,1:2]
site_covs <-as.data.frame(apply(site_covs,2,scale))
names(site_covs) <- c("elev", "dist_casa")


umf <-  unmarkedFrameOccuMulti(y=y, 
                              siteCovs=site_covs,
                              obsCovs = NULL)
plot(umf)
#umf

# occFormulas Length should match number/order of columns in fDesign
umf@fDesign


fit1 <- occuMulti(detformulas, stateformulas, umf,    
        method="BFGS", se=TRUE, engine=c("C"), silent=FALSE)

fit1

# update model
# occFormulas2 <- c('~dist_casa', '~dist_casa', '~dist_casa', "~1", "~1", "~1", "~0")
occFormulas2 <- c('~dist_casa', '~dist_casa', "~0")
fit2 <- update(fit1, stateformulas=occFormulas2)
fit2

#List of fitted models
fl <- fitList(fit1, fit2)
coef(fl)

#Model selection
modSel(fl)

bt <- parboot(fit1, nsim=50) # takes time best model
plot(bt)

```

### plot predicted marginal occupancy

```{r}
#Plot predicted marginal occupancy as a function of disturbance
r <- range(cams_ucu_sf$elev)
x1 <- seq(r[1],r[2],length.out=100)
x_scale <- (x1-mean(cams_ucu_sf$elev))/sd(cams_ucu_sf$elev)

r2 <- range(cams_ucu_sf$dist_casa)
x2 <- seq(r2[1],r2[2],length.out=100)
x2_scale <- (x2-mean(cams_ucu_sf$dist_casa))/sd(cams_ucu_sf$dist_casa)

nd <- matrix(NA, 100, 2)
nd <- data.frame(elev=x_scale, dist_casa= x2_scale)



tigrinus_pred <- predict(fit1, "state", species="tigrinus", newdata=nd)
tigrinus_pred$Species <- "tigrinus"

perros_pred <- predict(fit1, "state", species="perros", newdata=nd)
perros_pred$Species <- "perros"

# ocelote_pred <- predict(fit2, "state", species="ocelote", newdata=nd)
# ocelote_pred$Species <- "ocelote"


plot_dat <- rbind(tigrinus_pred, perros_pred)#, ocelote_pred)

ggplot(data=plot_dat, aes(x=rep(x1,2), y=Predicted)) + # change to 3 sp and x2 to distance
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Species), alpha=0.3) +
  geom_line(aes(col=Species)) +
  labs(x="Elevation", y="Marginal occupancy") +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        axis.text=element_text(size=12), axis.title=element_text(size=14),
        legend.text=element_text(size=12), legend.title=element_text(size=14))



```

### plot predicted co-occurrence occupancy

```{r}
#Plot predicted marginal occupancy as a function of disturbance
r <- range(cams_ucu_sf$elev)
x1 <- seq(r[1],r[2],length.out=100)
x_scale <- (x1-mean(cams_ucu_sf$elev))/sd(cams_ucu_sf$elev)

r2 <- range(cams_ucu_sf$dist_casa)
x2 <- seq(r2[1],r2[2],length.out=100)
x2_scale <- (x2-mean(cams_ucu_sf$dist_casa))/sd(cams_ucu_sf$dist_casa)

nd <- matrix(NA, 100, 2)
nd <- data.frame(elev=x_scale, dist_casa= x2_scale)


tigrinus_perros_pred <- predict(fit1, "state", 
                         species=c("tigrinus", "perros"), newdata=nd)

# tigrinus_pred$Species <- c("tigrinus", "perros")

# perros_pred <- predict(fit1, "state", species="perros", newdata=nd)
# perros_pred$Species <- "perros"

# ocelote_pred <- predict(fit2, "state", species="ocelote", newdata=nd)
# ocelote_pred$Species <- "ocelote"


plot_dat <- tigrinus_perros_pred #rbind(tigrinus_pred, perros_pred)#, ocelote_pred)

ggplot(data=plot_dat, aes(x=rep(x1), y=Predicted)) + # change to 3 sp and x2 to distance
  geom_ribbon(aes(ymin=lower, ymax=upper, fill = "grey50"), alpha=0.3) +
  geom_line(aes(y=Predicted), col="blue") +
  labs(x="Elevation", y="co-occurence") +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        axis.text=element_text(size=12), axis.title=element_text(size=14),
        legend.text=element_text(size=12), legend.title=element_text(size=14))



```


### plot predicted conditional occupancy 

```{r}
#Plot predicted marginal occupancy as a function of disturbance
r <- range(cams_ucu_sf$elev)
x1 <- seq(r[1],r[2],length.out=100)
x_scale <- (x1-mean(cams_ucu_sf$elev))/sd(cams_ucu_sf$elev)

r2 <- range(cams_ucu_sf$dist_casa)
x2 <- seq(r2[1],r2[2],length.out=100)
x2_scale <- (x2-mean(cams_ucu_sf$dist_casa))/sd(cams_ucu_sf$dist_casa)

nd <- matrix(NA, 100, 2)
nd <- data.frame(elev=x_scale, dist_casa= x2_scale)

##### conditional

tigrinus_perro_no <- predict(fit1, "state", 
                         species="tigrinus", 
                         cond='-perros',
                         newdata=nd)

tigrinus_perro_no$Species <- "perro ausente"

tigrinus_perro_si <- predict(fit1, "state", 
                         species="tigrinus", 
                         cond='perros',
                         newdata=nd)

tigrinus_perro_si$Species <- "perro presente"





# perros_pred <- predict(fit1, "state", species="perros", newdata=nd)
# perros_pred$Species <- "perros"

# ocelote_pred <- predict(fit2, "state", species="ocelote", newdata=nd)
# ocelote_pred$Species <- "ocelote"


plot_dat <- rbind(tigrinus_perro_si, tigrinus_perro_no)#, ocelote_pred)

ggplot(data=plot_dat, aes(x=rep(x1,2), y=Predicted)) + # change to 3 sp and x2 to distance
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Species), alpha=0.3) +
  geom_line(aes(col=Species)) +
  labs(x="Elevation", y="tigrinus conditional occupancy") +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        axis.text=element_text(size=12), axis.title=element_text(size=14),
        legend.text=element_text(size=12), legend.title=element_text(size=14))



```






# JUNTANDO TODOS LOS DATOS



```{r}
library(readxl)
total <- read_delim("C:/CodigoR/tigrinus2/data/total.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)


```

## crerar matrices para unmarked

```{r}
#funcion para crear todas las tablas de datos
all_data_total <-  f.matrix.creator2 (total)
# names(all_data_total) # ver lass especies y en que lista esta cada una
# Tigrinus es lista 7

datatable(
data = as.data.frame(names(all_data_total)),
caption = "Especies Ucumari",
filter = "top"
)



#code to shrink the matrix to exactly 150 columns: Aprox una semana
f.shrink.matrix.to150<-function(matrix){
  nc<-dim(matrix)[2]
  if(!nc%%150){ # of the number of columns is exactly divisible by 15
    newc<-nc%/%150
    old.cols<-seq(1,nc,newc)
    new.matrix<-matrix(NA,nr=nrow(matrix),nc=150)
    for(i in 1:150){
      new.matrix[,i]<-apply(matrix[,old.cols[i]:(old.cols[i]+newc-1)],1,max,na.rm=T)
    }
  } else{
    rem<-nc%%150
    newc<-nc%/%150
    old.cols<-seq(1,nc-rem,newc)
    new.matrix<-matrix(NA,nr=nrow(matrix),nc=150)
    for(i in 1:149)
      new.matrix[,i]<-apply(matrix[,old.cols[i]:(old.cols[i]+newc-1)],1,max,na.rm=T)
    new.matrix[,150]<-apply(matrix[,old.cols[150]:nc],1,max,na.rm=T) 
  }
  new.matrix[new.matrix=="-Inf"]<-NA
  rownames(new.matrix)<-rownames(matrix)
  new.matrix
}




# tabla con solo tiginus
tigrinus_total <- all_data_total[[7]]

# colapsa a una semana take tima
tigrinus_150<-f.shrink.matrix.to150(matrix = all_data_total[[7]])


# crear objeto umf
umf_tigrinus_total <- unmarkedFrameOccu(y=tigrinus_150)
# verificar datos en grafica
plot(umf_tigrinus_total)




```










# LAGUNA DE LA COCHA

```{r}

library(readr)
cocha1_data <- read_excel("C:/CodigoR/tigrinus2/data/Full_data_Ucumari_Huila_Cocha1_Cocha2.xlsx", 
    sheet = "cocha1")

cocha2_data <- read_excel("C:/CodigoR/tigrinus2/data/Full_data_Ucumari_Huila_Cocha1_Cocha2.xlsx", 
    sheet = "cocha2")

cocha2_data$binomial <- str_c (cocha2_data$Genus, "_", cocha2_data$Species)




```

## crerar matrices para unmarked

```{r}
#| column: screen-inset-shaded


############# start spatial part
#### make sf object
cocha1_sf <- st_as_sf(cocha1_data, coords = c("Longitude", "Latitude"), crs = "EPSG:4326")
#### make sf object
cocha2_sf <- st_as_sf(cocha2_data, coords = c("Longitude", "Latitude"), crs = "EPSG:4326")


# get centroid. 1st bbox, make pol
centroid_cocha <- st_centroid(st_as_sfc(st_bbox(cocha1_sf)))
# get altitude
elev_cocha_full <- elevation_3s(centroid_cocha[[1]][1], centroid_cocha[[1]][2], 
                         path="C:/CodigoR/tigrinus2/raster")

# elev_ucu_full_ras <- elev_ucu_full %>% raster()

# st_bbox(ucumari_sf) # notice order xmin, xmax, ymin, ymax
ext_cocha <- ext(-77.17, -77.05,  0.886,    1.085 )
elev_cocha <- crop(elev_cocha_full, ext_cocha) 

# convert from terra to raster
elev_cocha_ras <-  elev_cocha %>% raster()

##### get uniques
cams_cocha1 <-cocha1_data %>% dplyr::select(c("Longitude", "Latitude", "camera_trap")) %>% distinct()
#### make sf object
cams_cocha1_sf <- st_as_sf(cams_cocha1, coords = c("Longitude", "Latitude"), crs = "EPSG:4326")

cams_cocha2 <-cocha2_data %>% dplyr::select(c("Longitude", "Latitude", "camera_trap")) %>% distinct()
#### make sf object
cams_cocha2_sf <- st_as_sf(cams_cocha2, coords = c("Longitude", "Latitude"), crs = "EPSG:4326")



# extract values from raster using altitude
cams_cocha1_sf$elev <- extract(elev_cocha_ras, cams_cocha1_sf)
cams_cocha2_sf$elev <- extract(elev_cocha_ras, cams_cocha2_sf)


############## make distance map
# # Convert points to sp spatialpointdatafram
# casas_points <- as(casas_sf, "Spatial")
# # Projection
# # Be sure to have rgdal first installed.
# casas_points_utm <- spTransform(casas_points, CRS('+init=epsg:32718'))
# # convert to ppp
# casas_points_ppp <- as(as(casas_points_utm, "SpatialPoints"), "ppp")
# # distance
# casas_distance <- distmap(casas_points_ppp)

####### extract distance 
# 
# casas_distance_ras<- raster(casas_distance) # convert raster
# crs(casas_distance_ras) <- '+init=epsg:32718' # add crs
# 
# # project raster
# casa_distance <- projectRaster(casas_distance_ras, crs = crs(casas_points))


# extracr raster values
# cams_ucu_sf$dist_casa <- raster::extract(casas_distance_ras, cams_ucu_sf)
cams_cocha1_sf$dist_casa <- raster::extract(casa_distance, cams_cocha1_sf) # also works
cams_cocha2_sf$dist_casa <- raster::extract(casa_distance, cams_cocha2_sf) # also works


# plot map
mapview(elev_cocha_ras) + mapview(cams_cocha1_sf["camera_trap"]) + mapview(cams_cocha2_sf["camera_trap"]) 

############### end spatial part

# fix cocha2 data
# cocha2_data$binomial <- str_c (cocha2_data$Genus, "_", cocha2_data$Species)

#funcion para crear todas las tablas de datos
all_data_cocha1 <-  f.matrix.creator2 (cocha1_data)
all_data_cocha2 <-  f.matrix.creator2 (cocha2_data)
# names(all_data) # ver lass especies y en que lista esta cada una


# kable(names(all_data_cocha1)) # html table
# Tigrinus es lista 11 y perro 2

# kable(names(all_data_cocha2)) # html table
# Tigrinus es 30
# Perro no hay

datatable(
data = as.data.frame(names(all_data_cocha1)),
caption = "Especies Ucumari",
filter = "top"
)



```

## unmarked

```{r}
# tabla con solo tiginus
tigrinus_cocha1 <- all_data_cocha1[[11]]
tigrinus_cocha2 <- all_data_cocha2[[30]]
# cargar paquete
library(unmarked)
# crear objeto umf
umf_tigrinus_cocha1 <- unmarkedFrameOccu(y=tigrinus_cocha1)
umf_tigrinus_cocha2 <- unmarkedFrameOccu(y=tigrinus_cocha2)
# verificar datos en grafica
plot(umf_tigrinus_cocha1)
plot(umf_tigrinus_cocha2)


# tabla con solo perro
perros_cocha1 <- all_data_cocha1[[2]]
# cargar paquete
library(unmarked)
# crear objeto umf
umf_perros_cocha1 <- unmarkedFrameOccu(y=perros_cocha1)
# verificar datos en grafica
plot(umf_perros_cocha1)


```

## Modelo nulo

```{r}

# modelo nulo
fm_tig_cocha1 <- occu(~1 ~1, umf_tigrinus_cocha1)  # fit a model
fm_tig_cocha2 <- occu(~1 ~1, umf_tigrinus_cocha2)  # fit a model

backTransform(fm_tig_cocha1, type="det") #estimado lineal de deteccion
backTransform(fm_tig_cocha1, type="state") # estimado linel de ocupacion

backTransform(fm_tig_cocha2, type="det") #estimado lineal de deteccion
backTransform(fm_tig_cocha2, type="state") # estimado linel de ocupacion

# modelo nulo
fm_perros_cocha1 <- occu(~1 ~1, umf_perros_cocha1)  # fit a model

backTransform(fm_perros_cocha1, type="det") #estimado lineal de deteccion
backTransform(fm_perros_cocha1, type="state") # estimado linel de ocupacion


```


## MODELOS DE CO-OCURRENCIA

```{r}

detformulas <- c( "~1", "~1")#, "~1")
#stateformulas <- c('~elev','~elev', '~elev', "~1", "~1", "~1", "~0")# 3 sp
stateformulas <- c('~elev','~elev', "~0")
y <- list(tigrinus_cocha1, perros_cocha1)# , ocelote_ucu)
names(y) <- c("tigrinus", "perros")#, "ocelote")


obs_covs <-as.data.frame(scale(cams_cocha1_sf$dist_casa))
names(obs_covs) <- "dist_casa"

site_covs <- data.frame(cams_cocha1_sf[,c('elev','dist_casa')])[,1:2]
site_covs <-as.data.frame(apply(site_covs,2,scale))
names(site_covs) <- c("elev", "dist_casa")


umf <-  unmarkedFrameOccuMulti(y=y, 
                              siteCovs=site_covs,
                              obsCovs = NULL)
plot(umf)
#umf

# occFormulas Length should match number/order of columns in fDesign
umf@fDesign


fit1 <- occuMulti(detformulas, stateformulas, umf,    
        method="BFGS", se=TRUE, engine=c("C"), silent=FALSE)

fit1

# update model
# occFormulas2 <- c('~dist_casa', '~dist_casa', '~dist_casa', "~1", "~1", "~1", "~0")
occFormulas2 <- c('~dist_casa', '~dist_casa', "~0")
fit2 <- update(fit1, stateformulas=occFormulas2)
fit2

#List of fitted models
fl <- fitList(fit1, fit2)
coef(fl)

#Model selection
modSel(fl)

bt <- parboot(fit2, nsim=50) # takes time
plot(bt)

```

### plot predicted marginal occupancy

```{r}
#Plot predicted marginal occupancy as a function of disturbance
r <- range(cams_cocha1_sf$elev)
x1 <- seq(r[1],r[2],length.out=100)
x_scale <- (x1-mean(cams_cocha1_sf$elev))/sd(cams_cocha1_sf$elev)

r2 <- range(cams_cocha1_sf$dist_casa)
x2 <- seq(r2[1],r2[2],length.out=100)
x2_scale <- (x2-mean(cams_cocha1_sf$dist_casa))/sd(cams_cocha1_sf$dist_casa)

nd <- matrix(NA, 100, 2)
nd <- data.frame(elev=x_scale, dist_casa= x2_scale)



tigrinus_pred <- predict(fit2, "state", species="tigrinus", newdata=nd)
tigrinus_pred$Species <- "tigrinus"

perros_pred <- predict(fit2, "state", species="perros", newdata=nd)
perros_pred$Species <- "perros"

# ocelote_pred <- predict(fit2, "state", species="ocelote", newdata=nd)
# ocelote_pred$Species <- "ocelote"


plot_dat <- rbind(tigrinus_pred, perros_pred)#, ocelote_pred)

ggplot(data=plot_dat, aes(x=rep(x1,2), y=Predicted)) + # change to 3 sp and x2 to distance
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Species), alpha=0.3) +
  geom_line(aes(col=Species)) +
  labs(x="Elevation", y="Marginal occupancy") +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        axis.text=element_text(size=12), axis.title=element_text(size=14),
        legend.text=element_text(size=12), legend.title=element_text(size=14))



```


### plot predicted co-occurrence occupancy

```{r}
#Plot predicted marginal occupancy as a function of disturbance
r <- range(cams_cocha1_sf$elev)
x1 <- seq(r[1],r[2],length.out=100)
x_scale <- (x1-mean(cams_cocha1_sf$elev))/sd(cams_cocha1_sf$elev)

r2 <- range(cams_ucu_sf$dist_casa)
x2 <- seq(r2[1],r2[2],length.out=100)
x2_scale <- (x2-mean(cams_cocha1_sf$dist_casa))/sd(cams_cocha1_sf$dist_casa)

nd <- matrix(NA, 100, 2)
nd <- data.frame(elev=x_scale, dist_casa= x2_scale)


tigrinus_perros_pred <- predict(fit2, "state", 
                         species=c("tigrinus", "perros"), newdata=nd)

# tigrinus_pred$Species <- c("tigrinus", "perros")

# perros_pred <- predict(fit1, "state", species="perros", newdata=nd)
# perros_pred$Species <- "perros"

# ocelote_pred <- predict(fit2, "state", species="ocelote", newdata=nd)
# ocelote_pred$Species <- "ocelote"


plot_dat <- tigrinus_perros_pred #rbind(tigrinus_pred, perros_pred)#, ocelote_pred)

ggplot(data=plot_dat, aes(x=rep(x1), y=Predicted)) + # change to 3 sp and x2 to distance
  geom_ribbon(aes(ymin=lower, ymax=upper, fill = "grey50"), alpha=0.3) +
  geom_line(aes(y=Predicted), col="blue") +
  labs(x="Elevation", y="co-occurence") +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        axis.text=element_text(size=12), axis.title=element_text(size=14),
        legend.text=element_text(size=12), legend.title=element_text(size=14))



```


### plot predicted conditional occupancy 

```{r}
#Plot predicted marginal occupancy as a function of disturbance
r <- range(cams_cocha1_sf$elev)
x1 <- seq(r[1],r[2],length.out=100)
x_scale <- (x1-mean(cams_cocha1_sf$elev))/sd(cams_cocha1_sf$elev)

r2 <- range(cams_cocha1_sf$dist_casa)
x2 <- seq(r2[1],r2[2],length.out=100)
x2_scale <- (x2-mean(cams_cocha1_sf$dist_casa))/sd(cams_cocha1_sf$dist_casa)

nd <- matrix(NA, 100, 2)
nd <- data.frame(elev=x_scale, dist_casa= x2_scale)

##### conditional

tigrinus_perro_no <- predict(fit2, "state", 
                         species="tigrinus", 
                         cond='-perros',
                         newdata=nd)

tigrinus_perro_no$Species <- "perro ausente"

tigrinus_perro_si <- predict(fit2, "state", 
                         species="tigrinus", 
                         cond='perros',
                         newdata=nd)

tigrinus_perro_si$Species <- "perro presente"





# perros_pred <- predict(fit1, "state", species="perros", newdata=nd)
# perros_pred$Species <- "perros"

# ocelote_pred <- predict(fit2, "state", species="ocelote", newdata=nd)
# ocelote_pred$Species <- "ocelote"


plot_dat <- rbind(tigrinus_perro_si, tigrinus_perro_no)#, ocelote_pred)

ggplot(data=plot_dat, aes(x=rep(x1,2), y=Predicted)) + # change to 3 sp and x2 to distance
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Species), alpha=0.3) +
  geom_line(aes(col=Species)) +
  labs(x="Elevation", y="tigrinus conditional occupancy") +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        axis.text=element_text(size=12), axis.title=element_text(size=14),
        legend.text=element_text(size=12), legend.title=element_text(size=14))



```





